---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# featureselection Package for R

Feature Selection with Machine Learning Models

<!-- badges: start -->
[![R build status](https://github.com/UBC-MDS/feature-selection-r/workflows/R-CMD-check/badge.svg)](https://github.com/UBC-MDS/feature-selection-r/actions)
[![codecov](https://codecov.io/gh/UBC-MDS/feature-selection-r/branch/master/graph/badge.svg)](https://codecov.io/gh/UBC-MDS/feature-selection-r)
<!-- badges: end -->

## Overview:

If you have encountered a dataset with a myriad number of features, it could be very difficult to work with them all. Some features may not even be important or relevant and could even cause optimization bias. One approach to this problem is to select a subset of these features for your model. Feature selection will reduce complexity, reduce the time when training an algorithm, and improve the accuracy of your model (if we select them wisely). However, this is not a trivial task.

The  **featureselection** package for `R` can help you with this task. It is similar to its companion package [featureselection Package for Python](https://github.com/UBC-MDS/feature-selection-python).

## Features

In this package, four functions are included for feature selection:

* `forward_selection` - Function that use the forward selection algorithm to choose the number of features in a model. This iterative algorithm starts as an empty model, and add the feature with the best improvement based on the accuracy of the model. The process then is iteratively repeated selecting the features with the best improvement in accuracy. This procedure stops when the remaining features no longer enhance the accuracy of the model.

* `recursive_feature_elimination` - Iteratively fit and score an estimator for greedy feature elimination. The model initially considers all features with the goal of discovering the worst performing feature which is then removed from the dataset. This process is repeated until the desired number of features are attained.

* `variance_thresholding` - Perform simmulated annealing to select features: randomly choose a set of features and determine model performance. Then slightly modify the chosen features randomly and test to see if the modified feature list has improved model performance. If there is improvement, the newer model is kept, if not, a test is performed to determine if the worse model is still kept based on a acceptance probability that decreases as iterations continue and how worse the newer model performs. The process is repeated for a set number of iterations.

* `simulated_annealing` - Perform simmulated annealing to select features: randomly choose a set of features and determine model performance. Then slightly modify the chosen features randomly and test to see if the modified feature list has improved model performance. If there is improvement, the newer model is kept, if not, a test is performed to determine if the worse model is still kept based on a acceptance probability that decreases as iterations continue and how worse the newer model performs. The process is repeated for a set number of iterations.

## Existing Ecosystems:

Some of the above features already exsist within the R ecosystem but are provided for feature parity with [Python edition](https://github.com/UBC-MDS/feature-selection-python) of this package.

- [Forward Selection](https://www.rdocumentation.org/packages/MXM/versions/0.9.4/topics/Forward%20selection)

- [Recursive Feature Elimination](https://www.rdocumentation.org/packages/caret/versions/6.0-85/topics/rfe)

- Variance Threshold (None)

- Simulated Annealing (None)

## Installation:

Make sure you have the `devtools` package installed. You can install it as follows.

```{r eval=FALSE}
#Install development version from Github
install.packages("devtools")
```

Then, install the feature selection package.

```{r eval=FALSE}
devtools::install_github("UBC-MDS/feature-selection-r")
```

## Dependencies

- [R 3.6](https://www.r-project.org/)
- [dyplr 0.8.0](https://dplyr.tidyverse.org/)
- [purrr 0.3.0](https://purrr.tidyverse.org/)
- [stats 3.6.0](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html)

## Usage

To guide you with an example of how to use this package, we will use the [Friedman dataset](https://www.rdocumentation.org/packages/tgp/versions/2.4-14/topics/friedman.1.data).

Load dataset:
```{r eval=FALSE}
data <- dplyr::select(tgp::friedman.1.data(), -Ytrue)
X <- dplyr::select(data, -Y)
y <- dplyr::select(data, Y)
```

Use of feature selection functions:

### forward_selection

```{r eval=FALSE}
# Create a 'scorer' that accepts a dataset
# and returns the Mean Squared Error.
custom_scorer <- function(data){
  model <- lm(Y ~ ., data)
  return(mean(model$residuals^2))
}

# use function
featureselection::forward_selection(custom_scorer, X, y, 3, 7)
[1] 4 2 1 5
```

### recursive_feature_elimination

```{r eval=FALSE}
# Create a custom 'scorer' that accepts a dataset and returns
# the name of the column with the lowest coefficient weight.
custom_scorer <- function(data){
  model <- lm(Y ~ ., data)
  names(which.min(model$coefficients[-1]))[[1]]
}

# use function
featureselection::recursive_feature_elimination(custom_scorer, X, y, 4)
[1] "X1" "X2" "X4" "X5" "Y"
```

### simulated_annealing

```{r eval=FALSE}
# Create a 'scorer' that accepts a dataset
# and returns the Mean Squared Error.
custom_scorer <- function(data){
  model <- lm(Y ~ ., data)
  return(mean(model$residuals^2))
}

# use function
featureselection::simulated_annealing(custom_scorer, X, y)
[1]  1  2  3  4  5  7  9 10
```

### variance_threshold_select

*Note: This function does not use the Friedman dataset.*
```
# sample data to test variance
data <- data.frame(x1=c(1,2,3,4,5), x2=c(0,0,0,0,0), x3=c(1,1,1,1,1))

# use function
featureselection::variance_threshold_select(data)
[1] 1
```

## Documentation
The official documentation is hosted on Read the Docs: <https://feature-selection.readthedocs.io/en/latest/>

## Credits
This package was created with Cookiecutter and the UBC-MDS/cookiecutter-ubc-mds project template, modified from the [pyOpenSci/cookiecutter-pyopensci](https://github.com/pyOpenSci/cookiecutter-pyopensci) project template and the [audreyr/cookiecutter-pypackage](https://github.com/audreyr/cookiecutter-pypackage).
